A Technique for Inter-Catalogue Recommendations

### 2014-06-11 paper2/paper.tex
Modify Latent Concept section

\subsection*{Leveraging Recommendations}
This $ContentBased Recommender$ is limited in a few ways. For one, it attempts
to capture similarity between products by looking exclusively at the words in
their descriptions. One issue here is that brands tend to use different 
vocabularies to describe the their products. One may refer to a color as 
``off-white'' while another calls the same color ``ivory.'' To the $ContentBased
Recommender$, these words are as different as ``black'' and ``white.'' Another 
issue is that the best recommendation for a product may not be the most similar
other product. For instance, a striped black-and-white button-up shirt could be 
an excellent recommendation for a blue checkered button-up, even though the two 
have ostensibly different descriptions.

We can alleviate some of these issues by leveraging the recommendations we
already have between products within the same catalogue. One approach is to 
consider the {\emph neighborhood} of each product, that is that product along 
with each of its neighbors in the recommendation graph. With this we create a
new similarity function, $NeighborhoodSimilarity$. This function uses the 
$ContentSimilarity$ function from before to compare every pair of products 
between the two neighborhoods:

@@

\subsection*{Latent Concepts}
Although the $Neighborhood Recommender$ alleviates some of the issues of 
content-based similarity, it is still fundamentally based on the descriptions of
products. Ideally, we would like to recommend products that are similar in more
fundamental ways. If a dress is elegant and modern, we would like to recommend 
another dress that is elegant and modern, even if those words are not explicitly
in the descriptions of either. These types of underlying concepts can be 
intangible and difficult to define. However, if we assume that the given
recommendations are truly based on such latent concepts, we can attempt to learn 
them by analyzing the graphs.

One common approach to learning latent semantic factors is to apply Singular Value
Decomposition to . In
traditional LSI, used for Natural Language Processing, concepts are learned over 
a dataset consists of documents, each containing terms. Singular Value 
Decomposition (SVD) is applied to the term-document matrix and the resulting 
three matrices are truncated to the number of latent concepts desired. The 
resulting left matrix is a term-concept matrix, where each row is a term 
represented in concept space.

We apply the same idea to the product catalogues in order to learn the latent
concepts of each product. Analogous to documents containing terms, we look at
user sessions which contain the products viewed by that user. Although we do not
have actual user data, we can simulate user sessions by traversing the
recommendation graphs.

To simulate user sessions, we use a random walk technique on the graph. We
assume that the edges of a recommendation graph indeed connect products that a user
would co-view. Thus traversing the edges of a graph is a reasonable simulation
of a user session. For details on these random walks, see the section titled
Random Walks in the Implementation section below.

With the simulated user sessions, we can implement LSI on each of the catalogues
and learn the latent concepts for each product. We want to find the products from
one catalogue that are most similar in latent concept space to products in the
other catalogue. However, since the user sessions (random walks) include
products either from one catalogue or the other, we learn different latent 
spaces for each catalogue. The problem is therefore reduced to finding a mapping
from one concept space to the other. If this can be accomplished, we
can map the concepts of each product into concepts of the other catalogue. This
creates a concept representation in the other catalouge's concept space, meaning
that we can then find nearest neighbors in that catalogue's concept space.

To map concepts from one catalogue to another, we must revisit
content-similarity. In this case, we represent each concept as a vector of
words. To calculate the value for a word w for concept c, we sum the tf-idf
value of w across all products, weighted by how much that product is represented
by concept c. Thus if a concept could be described as formal, the words
``sophisticated'' or ``official'' may have high value since products with those
descriptive words would also have a high amount of the formal concept in their
concept representations. Using these vector representation for concepts, we can
simply calculate the dot product between two concepts to find how similar they
are, in a relative sense. This can be used to create a mapping from one set of
concepts to another.

In all, the LSI approach consists of four phases. First we simulate user
sessions by randomly walking the recommendation graphs. Second we learn concept
spaces for products in each catalogue via SVD. Third we find a mapping between
concept spaces. And lastly we find the nearest neighbors of each product in the
other catalogue's concept space. These nearest neighbors are the predicted
recommendations between products in different catalogues.

@@

\subsection*{Concept Mapping}
A tricky operation of this procedure is the concept mapping. We attempt to find
the best linear transformation from one concept space to the other. We do this
by representing each concept as a vector of words in the vocabulary, each
weighted with how well it represents the concept. We construct this by weighing
the tfidf of each word in a product's description by how well that word
represents the concept. Note that products can have a positive or negative value
for each concept, thus the word representation of concpets are also positive and
negative.

@@

One way to normalize is to normalize each column of the inner-product matrix by
the L1 norm. This preserves the L1 norm of products after the transformation,
but does not preserve the L2 norm. Similarly, we can normalize the columns by
the L2 norm. Neither of these methods preserve the L2 norm, which would be ideal
as we find the nearest neighbors in the other concept space by their L2
distance.

@@

Proceeding as before, we find the nearest products in concept space. However,
before deeming the nearest neighbors to be the best recommendations, we divide
the distance to each product by its popularity. Since we model distance in 
concept space to be inversely proportional to similarity, this reincorporation
of popularity is analagous to multiplying similarity by popularity, as done in
the original construction of the recommendation graph.

@@

It is also informative to test the efficacy of the concept mapping, since this
such a hard task. As in product-product content-similarity, it is difficult to
capture similarity using descriptions alone. It is not even clear that a mapping 
necessarily exists between two seperately learned concept spaces. The concepts 
may represent fundamentally different things, which cannot be well mapped to 
each other. We can measure the efficacy of the mapping by comparing it to a 
couple of baselines.

The first baseline is in fact an upper bound of sorts on the mapping. This
baseline recommender is called the One Model Recommender. Unlike the other
baselines, this recommender is not a valid one, as it cheats by using
information from the unpartitioned graph. The One Model Recommender learns a
latent concept space over the entire recommendation graph by allowing the random
walkers to traverse it before it is partitioned. In every other respect, it is
the same as the LSI Recommender. However, since products from both catalogues
are represented in the same concept space, there is no need for a mapping
between concept spaces.

@@

\subsection*{Results}
[Number of products mapped to in different methods]
\begin{center}
\begin{tabular}{ | l | c | c | c |}
\hline
Normalization & 2-Precision & 2-Recall & 2-F1 Score \\ \hline\hline
L1 normalize each column of matrix &&&\\ \hline
L2 normalize each column of matrix &&&\\ \hline
Orthogonalize matrix &&&\\ \hline
Normalize products in concept space &&&\\ \hline
\end{tabular}
\end{center}

>>>>

\subsection*{Leveraging Recommendations}
This $ContentBased Recommender$ is limited in a few ways. For one, it attempts
to capture similarity between products by looking exclusively at the words in
their descriptions. One issue here is that brands tend to use different 
vocabularies to describe the their products. One may refer to a color as 
``off-white'' while another calls the same color ``ivory.'' To the $ContentBased
Recommender$, these words are as different as ``black'' and ``white.'' Another 
issue is that the best recommendation for a product may not be the most similar
product. For instance, a striped black-and-white button-up shirt could be 
an excellent recommendation for a blue checkered button-up, even though the two 
have ostensibly different descriptions.

We can alleviate some of these issues by leveraging the recommendations we
already have between products within the same catalogue. One approach is to 
consider the {\em neighborhood} of each product, that is a product along 
with each of its neighbors in the recommendation graph. With this, we create a
new similarity function called $NeighborhoodSimilarity$. This function uses the 
$ContentSimilarity$ function from before to compare every pair of products 
between the two neighborhoods:

@@

\subsection*{Latent Concepts}
Although the $Neighborhood Recommender$ alleviates some of the issues of 
content-based similarity, it is still primarily based on the descriptions of
products. Ideally, we would like to recommend products that are similar in more
fundamental ways. If a dress is elegant and modern, we would like to recommend 
another dress that is elegant and modern, even if those words are not explicitly
in the descriptions of either. These types of underlying concepts can be 
intangible and difficult to define. However, if we assume that the given
recommendations are truly based on such latent concepts, we can attempt to learn 
them by analyzing the graphs.

One common approach to learning latent semantic factors is to apply {\em
Singular Value Decomposition} (SVD) to a matrix of user data, as in many 
collaborative filtering methods [CITE]. This method requires a product-user
matrix, where each column represents the products viewed in a user session.
After SVD is applied to the matrix, the resulting three matrices are truncated
to the number of latent concepts desired. The resulting left matrix becomes a
product-concept matrix, where each row is a product represented in 
{\em concept-space}

In this experiment, there is no such product-user matrix to apply SVD.
However, since the recommendation graphs are constructed from user behavior,
we can use them to simulate user sessions. To do this, we use a random walk 
technique on the graph. We assume that the edges of a recommendation graph 
indeed connect products that a user would co-view. Thus traversing the edges 
of a graph is a reasonable simulation of a user session. For details on these
random walks, see the section titled Random Walks in the Implementation section
below.

With simulated user data for each recommendation graph, we can apply SVD on the
matrices from the two catalogues to learn the latent concepts for each product.
We want to find the products from one catalogue that are most similar in latent
concept-space to products in the other catalogue. However, since each user
session/random walk includes products exclusively from one of the catalogues, we
learn different latent spaces for each catalogue. Note that this would be an
issue even if we had the actual user data used to construct the two
recommendation graphs. The problem therefore becomes finding a mapping from one
concept-space to the other. If this can be accomplished, we can represent each
product of one catalogue in terms of concepts of the other catalogue and then
find the nearest neighbors of that catalogue by distance in concept-space.

To map concepts from one catalogue to another, we must revisit
content-similarity. In this case, we represent each concept (rather than each
product) as a vector of words. To represent a concept in terms of the
vocabulary, we calculate the value of each word by summing the
tf-idf of that word for each product $P$, weighted by the projection of $P$ on
our concept in concept-space. That is:

\begin{align}
C_w = \sum\limits_{P} \text{tfidf}(w)\times P_C
\end{align}
where $C_w$ is the value of concept $C$'s representation in dimension $w$, the sum 
is over all products $P$ in the catalgoue, tfidf($w$) is the tf-idf score for
word $w$, and $P_C$ is the projection of $P$ on concpet $C$'s axis in
concept-space.

Now for a concept such as {\em formal}, the words ``sophisticated'' or
``official'' may have high value since products with those descriptive words
would also have a large amount of the {\em formal} concept in their
concept-representations. Using these vector representation, we calculate the
inner product between two concepts to find how similar they are to eachother.
This can be used to create a mapping from one set of concepts to another as it
captures the relative breakdown of a concept from one space to the concepts in the
other.

In all, the latent-concept approach consists of four phases. First, we simulate user
sessions by randomly walking the recommendation graphs. Second, we learn concept
spaces for products in each catalogue via SVD. Third, we find a mapping between
concept-spaces using the aggregated descriptions of all products. And lastly, we
find the nearest neighbors of each product in the concept-space of the other 
catalogue concept-space. These nearest neighbors become the predicted
recommendations between products across the two catalogues. We call this method
the $LatentConcept Recommender$

@@

\subsection*{Concept Mapping}
A tricky operation of this procedure is the concept mapping. We attempt to find
the best linear transformation from one concept-space to the other. We do this
by representing each concept as a vector of words in the vocabulary, each
weighted with how well it represents the concept. We construct this by weighing
the tfidf of each word in a product's description by how well that word
represents the concept. Note that products can have a positive or negative value
for each concept, thus the word representation of concpets are also positive and
negative.

@@

One way to normalize is to normalize each column of the inner-product matrix by
the L1 norm. This preserves the L1 norm of products after the transformation,
but does not preserve the L2 norm. Similarly, we can normalize the columns by
the L2 norm. Neither of these methods preserve the L2 norm, which would be ideal
as we find the nearest neighbors in the other concept-space by their L2
distance.

@@

Proceeding as before, we find the nearest products in concept-space. However,
before deeming the nearest neighbors to be the best recommendations, we divide
the distance to each product by its popularity. Since we model distance in 
concept-space to be inversely proportional to similarity, this reincorporation
of popularity is analagous to multiplying similarity by popularity, as done in
the original construction of the recommendation graph.

@@

It is also informative to test the efficacy of the concept mapping, since this
such a hard task. As in product-product content-similarity, it is difficult to
capture similarity using descriptions alone. It is not even clear that a mapping 
necessarily exists between two seperately learned concept-spaces. The concepts 
may represent fundamentally different things, which cannot be well mapped to 
each other. We can measure the efficacy of the mapping by comparing it to a 
couple of baselines.

The first baseline is in fact an upper bound of sorts on the mapping. This
baseline recommender is called the One Model Recommender. Unlike the other
baselines, this recommender is not a valid one, as it cheats by using
information from the unpartitioned graph. The One Model Recommender learns a
latent concept-space over the entire recommendation graph by allowing the random
walkers to traverse it before it is partitioned. In every other respect, it is
the same as the LSI Recommender. However, since products from both catalogues
are represented in the same concept-space, there is no need for a mapping
between concept-spaces.

@@

\subsection*{Results}
[Number of products mapped to in different methods]
\begin{center}
\begin{tabular}{ | l | c | c | c |}
\hline
Normalization & 2-Precision & 2-Recall & 2-F1 Score \\ \hline\hline
L1 normalize each column of matrix &&&\\ \hline
L2 normalize each column of matrix &&&\\ \hline
Orthogonalize matrix &&&\\ \hline
Normalize products in concept-space &&&\\ \hline
\end{tabular}
\end{center}
